---
title: "PA1_template"
output: html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='figure/')
```
##Loading the data

First we load the data and required packages.  

```{r, echo=TRUE}
require(plyr)
require(ggplot2)
rawdata<-read.csv("activity.csv")
head(rawdata)
```

  
##What is mean total number of steps taken per day?  

###Make a histogram of the total number of steps taken each day

```{r Q1_1, echo=TRUE}
processed<-ddply(rawdata,.(date),summarize,steps=sum(steps,na.rm=TRUE))
head(processed)

hist(processed$steps,xlab="Steps per Day", main=paste("Steps Taken Per Day"))

```

###Calculate and report the mean and median total number of steps taken per day
```{r Q1_2, echo=TRUE}
meansteps<- mean(processed$steps,na.rm= TRUE)
mediansteps<-median(processed$steps,na.rm= TRUE)
```

The mean number of steps taken each day is `r round(meansteps)` and the median is `r round(mediansteps)`.  

  
##What is the average daily activity pattern?  

###Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r Q2_1, echo=TRUE}
processedts<-ddply(rawdata,.(interval),summarise,steps=mean(steps,na.rm=TRUE))
plot(processedts$interval, processedts$steps, type="l", xlab="Interval",ylab="Average Steps Taken", main="Average Steps Taken per 5 minute Interval")
maxinterval <-processedts$interval[which.max(processedts$steps)]
```

###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
The interval with the most steps taken on average is `r maxinterval`  

  
##Imputing Missing Values  

###Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r Q3_1, echo=TRUE}
numna<- sum(is.na(rawdata$steps))

```
Number of missing values = `r numna`

###Devise a strategy for filling in all of the missing values in the dataset. + Create a new dataset that is equal to the original dataset but with the missing data filled in.
Using mean for the 5-minute interval to impute NAs  
1. Pull out rows with NA  
2. Join using previous dataset  
3. Add it back to the original data set while removing NA rows
```{r Q3_2, echo=TRUE}
moddata<-rawdata[is.na(rawdata$steps),]
moddata[,1]=NULL
moddata<-join(moddata, processedts, by="interval",type="left")
moddata<-rbind(rawdata,moddata)
moddata<-na.omit(moddata)
```

###Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r Q3_3, echo=TRUE}
processedmod<-ddply(moddata,.(date),summarize,steps=sum(steps,na.rm=TRUE))
hist(processedmod$steps,xlab="Steps per Day", main=paste("Steps Taken Per Day"))

meanstepsmod<- mean(processedmod$steps)
medianstepsmod<-median(processedmod$steps)

```

The mean number of steps taken each day is `r round(meanstepsmod)` and the median is `r round(medianstepsmod)`.  
Imputing missing data raises the mean and median total daily number of steps by `r round(meanstepsmod - meansteps)` and `r round(medianstepsmod - mediansteps)` (rounded) respectively.  

##Are there differences in activity patterns between weekdays and weekends?  

###Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.  

```{r Q4_1, echo=TRUE}
rawdata$day<-weekdays(as.Date(rawdata$date))
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
rawdata$weekday <-  factor((rawdata$day %in% weekdays1)+1L, levels=1:2, labels=c('weekend', 'weekday'))

head(rawdata)
```


###Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).  


```{r Q4_2, echo=TRUE, fig.width=10, fig.height=12}
processedweekday<-ddply(rawdata,.(weekday, interval),summarise,steps=mean(steps,na.rm=TRUE))
par(mar=c(5,15,4,1))
par( mfrow = c( 2, 1 ) )
with(subset(processedweekday,weekday=="weekday"),plot(interval, steps, type="l", xlab="Interval",ylab="Average Steps Taken", main="Average Steps Taken per 5 minute Interval (Weekday)"))
with(subset(processedweekday,weekday=="weekend"),plot(interval, steps, type="l", xlab="Interval",ylab="Average Steps Taken", main="Average Steps Taken per 5 minute Interval (Weekend)"))
```